MS-4.2 Types of Relationships

### 2020-05-17 ms-a4-2_non-linear-relationships.rmd
Updated syllabus and started adding lesson 1

---
title: "MS-4.2 Types of Relationships"
output:
  html_document:
    code_folding: "hide"
---

@@

* use an exponential model to solve problems
  + graph and recognise an exponential function in the form $y=a^x$ and $y=a^{-x} (a>0)$ using technology
  + interpret the meaning of the intercepts of an exponential graph in a variety of contexts
  + construct and analyse an exponential model to solve a practical growth or decay problem
* construct and analyse a quadratic model to solve practical problems involving quadratic functions of expressions of the form $y=ax^2+bx+c$, for example, braking distance against speed
  + recognise the shape of a parabola and that it always has a turning point and an axis of symmetry
  + graph a quadratic function using technology
  + interpret the turning point and intercepts of a parabola in a practical context
  + consider the range of values for $x$ and $y$ for which the quadratic model makes sense in ap ractical context
* recognise that reciprocal functions of the form $y=\frac{k}{x}$, where $k$ is a constant, represent inverse variation, identify the rectangular hyperbolic shape of these graphs and their important features
  + use a reciprocal model to solve practical inverse variation problems algebraically and graphically, eg the amount of pizza received when sharing a pizza between increasing numbers of people
:::

@@

#  {.tabset .tabset-fade .tabset-pills}

## Exponentials {.tabset .tabset-fade .tabset-pills}

### Functions

```{r}
library("ggplot2")
eq = function(x){2^x}
ggplot(data.frame(x=c(-3, 3)), aes(x=x)) + 
  stat_function(fun=eq)
```

### Modelling

## Quadratics {.tabset .tabset-fade .tabset-pills}

### Functions

### Modelling

## Reciprocals {.tabset .tabset-fade .tabset-pills}

### Functions

### Modelling

## Inverse variation

## Miscellaneous problems

>>>>

---
title: "MS-4.2 Types of Relationships"
output:
  html_document:
    toc: true
    toc_float: true
---

@@

* use an exponential model to solve problems
  + graph and recognise an exponential function in the form $y=a^x$ and $y=a^{-x} (a>0)$ with and without technology
  + interpret the meaning of the intercepts of an exponential graph in a variety of contexts
  + construct and analyse an exponential model of the form $y=ka^x$ and $y=ka^{-x} (a>0)$ where $k$ is a constant, to solve a practical growth or decay problem
* construct and analyse a quadratic model to solve practical problems involving quadratic functions or expressions of the form $y=ax^2+bx+c$, for example, braking distance against speed
  + recognise the shape of a parabola and that it always has a turning point and an axis of symmetry
  + graph a quadratic function using technology
  + interpret the turning point and intercepts of a parabola in a practical context
  + consider the range of values for $x$ and $y$ for which the quadratic model makes sense in a practical context
* recognise that reciprocal functions of the form $y=\frac{k}{x}$, where $k$ is a constant, represent inverse variation, identify the rectangular hyperbolic shape of these graphs and their important features
  + use a reciprocal model to solve practical inverse variation problems algebraically and graphically, eg the amount of pizza received when sharing a pizza between increasing numbers of people
:::

# Exponential Functions [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_a_exponential-functions.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

@@

### Introduction

::: {.definition}
Exponential functions are functions that have the $x$ as the power of the constant.

i.e. $$y=ka^x$$ Where $a$ and $k$ are constants.
:::

For example, $$y=3\times2^x$$

::: {.note}
Note that the index operation happens before the mulitplication so $y=3\times2^x\neq6^x$
:::

### Graphs - Still sorting it out {.tabset .tabset-fade .tabset-pills}

#### $y=2^x$

#### $y=a^x$

#### $y=a^{-x}$

#### $y=a^x$ and $y=a^{-x}$

#### $y=ka^x$

### Observations

* The graph is above the $x$-axis.
  + $y$ can never be 0 or negative, unless $k<0$.
* The graph always passes through the point $(0, k)$. $(k\times a^0=k\times1=k)$
* Increasing the value of $a$ increases the steepness of the curve.
* Increasing the value of $x$ by 1 magnifies the value of $y$ by a factor of $a$
* The $x$-axis is an _asymptote_.
  + The curve never touches the $x$-axis.
* $a^x$ and $a^{-x}$ are reflections in the $y$-axis.
* $y=a^{-x}$ has the same graph as $y=\left(\dfrac{1}{a}\right)^x$
* $k$ stretches the graph vertically.

### Sketching

When sketching exponential graphs:

1. Construct a table of values
2. Draw a number plane
3. Plot at least 4 points
4. Draw a curve that goes through the points

### Examples {.tabset .tabset-fade .tabset-pills}

#### Example 1 {.tabset .tabset-fade .tabset-pills}

##### Question

Draw the graph of $y=3^x$

##### Show Solution

Embed

#### Example 2 {.tabset .tabset-fade .tabset-pills}

##### Question

Draw the graph of $y=3^{-x}$

##### Show Solution

Embed

# Exponential Models [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_b_exponential-models.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

# Quadratic Functions [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_c_quadratic-functions.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

# Quadratic Models [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_d_quadratic-models.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

# Reciprocal Functions [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_e_reciprocal-functions.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

# Reciprocal Models [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_f_reciprocal-models.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

# Inverse Variation [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_g_inverse-variation.html){target="blank"} {.tabset .tabset-fade .tabset-pills}

# Miscellaneous Problems [![](pres.png)](https://claytonpollard.github.io/stage6_standard_presentations/MS-A4.2_h_miscellaneous-problems.html){target="blank"} {.tabset .tabset-fade .tabset-pills}
